---
title: "Vignette ecophylo"
author: "Team bouclé"
date: "9 janvier 2021"
output: html_document
---

```{r setup, include=FALSE}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)
use_python(Sys.which("python3"))
```

## Rules for working in Rmarkdown with python.

*we will need to check the setup chunk for location of python3 binary across different os (ubuntu work with this code)*

### Python chunks calling R obj.

R chunks works just fine as usual.

```{r}
a <- 42 # set a in R
```

This chunk only work when the markdown is knitted.
See that we can call R object when looking for them in 'r' namespace.

```{python}
print(r.a)
r.a += 1 # modify R variable in python
print(r.a)
b = 666 # set b in python
```

### R chunks calling Python obj.

See that we can call python object when looking for them in 'py' list.

```{r}
print(a) # call a in R after modification in python
print(py$b) # call b in R
```

You can run python code in interactive mod in the console using the following 
function : *repl_python(quiet = T)*.
It set the console in 'python mode' so you can type in python commands. 
However you can't send lines from a chunk or a script.

To exit the consol in python, just type *exit* and tadaa the consol is back 
to R.


