---
title: "Vignette ecophylo"
author: "Team bouclé"
date: "9 janvier 2021"
output:
  html_document: default
  pdf_document: default
  
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)
use_python(Sys.which("python3"))
```

## Rules for working in Rmarkdown with python.

*we will need to check the setup chunk for location of python3 binary across different os (ubuntu work with this code)*

### Python chunks calling R obj.

R chunks works just fine as usual.

```{r}
py_config()
a <- 42 # set a in R
```

This chunk only work when the markdown is knitted.
See that we can call R object when looking for them in 'r' namespace.

```{python}
print(r.a)
r.a += 1 # modify R variable in python
print(r.a)
b = 666 # set b in python
```

### R chunks calling Python obj.

See that we can call python object when looking for them in 'py' list.

```{r}
print(a) # call a in R after modification in python
print(py$b) # call b in R
# Also possible to call directly python function in r
rand <- import('numpy')$random # import python module numpy and submod random
rand$randint(3) # call of function randint
```

You can run python code in interactive mod in the console using the following 
function : *repl_python(quiet = T)*.
It set the console in 'python mode' so you can type in python commands. 
However you can't send lines from a chunk or a script.

To exit the consol in python, just type *exit* and tadaa the consol is back 
to R.

### Using Python modules in R.

```{r, eval=FALSE}
# Be carefull for idiotproof of ecophylo and int class !
eco <- import('ecophylo')
eco$timeframes(I=as.integer(3), T=2, a=0.3)
# make dataframes with pandas is same as R
pd <- import('pandas')
d <- list(col1 = c(1,2,3), col2 = c(4,5,6))
pd$DataFrame(d)
```

### Simulating sub-populations linked by vicariance and migration events

```{r}
split_dates <- list(as.integer(100))

npop <- as.integer(2)
migrfrom <- as.integer(1.00)
migrto <- as.integer(2.00)
init_sizes <- list(rep(as.integer(500),2))
init_rates <- list(rep(as.integer(1),2))
m <- as.integer(0)

tree2 <- ecophylo$simulate(n, Jm, mu, 
                           split_dates,
                           npop,
                           init_rates,
                           init_sizes,
                           migrfrom,
                           migrto,
                           m,
                          seed = as.integer(42))

abund12 <- ecophylo$getAbund(tree2, n)
phylo2 <- read.tree(text = tree2$write())
```

## Intensive simulations along priors




```{r, eval=FALSE}
# NOT RUN
truc <- eco$dosimuls(nsim = 1, sample_size = 100, comprior = list(1000,10e9),
             muprior = list(1e-3) , verbose = TRUE, seed = 42)
```

```{python, eval=FALSE}
# NOT RUN
eco.dosimuls(nsim = 5, sample_size = 100, comprior = [1000,10e9], 
             muprior = [1e-3] , verbose = True, seed = 42)
```


<!--
## Adding Migrations
-->

