---
title: "Vignette ecophylo"
author: "Team bouclé"
date: "9 janvier 2021"
output:
  html_document: default
  pdf_document: default
  
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
library(reticulate)
knitr::opts_chunk$set(echo = TRUE)
use_python(Sys.which("python3"))
```

## Rules for working in Rmarkdown with python.

*we will need to check the setup chunk for location of python3 binary across different os (ubuntu work with this code)*

### Python chunks calling R obj.

R chunks works just fine as usual.

```{r}
py_config()
a <- 42 # set a in R
```

This chunk only work when the markdown is knitted.
See that we can call R object when looking for them in 'r' namespace.

```{python}
print(r.a)
r.a += 1 # modify R variable in python
print(r.a)
b = 666 # set b in python
```

### R chunks calling Python obj.

See that we can call python object when looking for them in 'py' list.

```{r}
print(a) # call a in R after modification in python
print(py$b) # call b in R
# Also possible to call directly python function in r
rand <- import('numpy')$random # import python module numpy and submod random
rand$randint(3) # call of function randint
```

You can run python code in interactive mod in the console using the following 
function : *repl_python(quiet = T)*.
It set the console in 'python mode' so you can type in python commands. 
However you can't send lines from a chunk or a script.

To exit the consol in python, just type *exit* and tadaa the consol is back 
to R.

### Using Python modules in R.

```{r, eval=FALSE}
# Be carefull for idiotproof of ecophylo and int class !
eco <- import('ecophylo')
eco$timeframes(I=as.integer(3), T=2, a=0.3)
# make dataframes with pandas is same as R
pd <- import('pandas')
d <- list(col1 = c(1,2,3), col2 = c(4,5,6))
pd$DataFrame(d)
```

### Coding a python package

To add code to the package function, once the files saved you need to recompile and reinstall the package.

This is done with the `dev/compil.sh` script, but only work if you have the write to install stuff on your computer.

After reinstalling the package, you need to restarte you R session so reticulate can find the new version of the package. This can be done with this R command :  

```{r, eval = FALSE}
.rs.restartR()
```

### Getting the prints of a python function in R

To gather the prints of a python function called from R, you need to wrap it into a reticulate function like below :

```{python}
def test(a):
  print(a)
  return 42
  
test(1)
```

```{r}
p = py_capture_output( x <- py$test(2) ) # wrap
cat(p) # the print output
print(x) # the return of the test function
```


